A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2019-08-29 rr-flow.Rmd
add further references close #15

In this tutorial, we describe a principled workflow to ensure long-term reproducibility of
R-based data analyses. In this context, reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and sustainably available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported. Here, we combine four tools from the computer sciences, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses. The major four elements are:

@@

The scientific community widely recognizes that these tools enhance reproducibility, and provides guidelines for their use (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018). However, only a workflow using them in unison allows confidence in reproducing a scientific report. Within the community of users of the R Programming Language for each purpose, a specific software solution is the most popular. For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`. Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher. Each of these software solutions is highly flexible and powerful, so to master them thoroughly often years of practice are required. Luckily the researcher can learn the minimal valuable subset needed to ensure reproducibility with little practice.

@@

@@

How to best share a reproducible workflow? We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017] or PsychData (citation needed), that is independ of Gihub, Docker Hub & MRAN. Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format. Typically, there is currently no single license that works for code, data, and media (such as text or figures). We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and moifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place. In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, rmarkdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license. Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work. These licenses are also in line with the recommendations by the Reproducibile Research Standard (see @stoddenEnhancingReproducibilityComputational2016).

>>>>

In this tutorial, we describe a principled workflow to ensure long-term reproducibility of
R-based data analyses. In this context, reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and sustainably available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992]. Here, we combine four tools from the computer sciences, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses. The major four elements are:

@@

The scientific community widely recognizes that these tools enhance reproducibility [@piccoloToolsTechniquesComputational2016], and provides guidelines for their use (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018). However, only a workflow using them in unison allows confidence in reproducing a scientific report. Within the community of users of the R Programming Language for each purpose, a specific software solution is the most popular. For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`. Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher. Each of these software solutions is highly flexible and powerful, so to master them thoroughly often years of practice are required. Luckily the researcher can learn the minimal valuable subset needed to ensure reproducibility with little practice.

@@

However, beside these tools that ease the process of creating a workflow on the researchers side, we see increased interest in changing how research is published and used[@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continues integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017, @yenniDevelopingModernData2018]. For such further technical advancements reproducibility is key and researcher using this workflow should be well prepared for such era of interactive publication.

@@

How to best share a reproducible workflow? We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017] or PsychData (citation needed), that is independ of Gihub, Docker Hub & MRAN. Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format. Typically, there is currently no single license that works for code, data, and media (such as text or figures). We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and moifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place. In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, rmarkdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license. Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work. These licenses are also in line with the recommendations by the Reproducibile Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
